---
title: "hw 6"
author: "rebecca park"
date: "11/5/2020"
output: pdf_document
---

## 1. Age restricted video games

```{r}
library(readxl)
VideoGames <- read_excel("~/Desktop/UGA/Fall 2020/STAT 6315/hw/hw 6/VideoGames.xls", n_max = 40)

# 1-way ANOVA
m1 = lm(Desire ~ Label, data = VideoGames)
anova(m1)

# critical F-value
qf(.95, 3, 36)
```
F-statistic = 1.8029
Degrees of freedom = 3
P-value = 0.1641

a. **No significant difference between age labels** - the calculated F-statistic is less than 2.866, the critical F-value. The p-value indicates there is a 0.164 probability of observing an F ratio greater than that given the null hypothesis is true, which is not quite a significant value. Thus accept the null hypothesis that there is no difference in rating based on the label. 

b. N/A - do not reject the null hypothesis.

## 2. Fatigue time
```{r}
# create data frame
class <- c(rep("normal", 7), rep("low_ow", 7), rep("med_ow", 7), rep("high_ow", 7))
time <- c(25, 28, 19, 27, 23, 30, 35, 
          24, 26, 18, 16, 14, 12, 17, 
          15, 18, 17, 25, 12, 10, 23, 
          10, 9, 18, 14, 6, 4, 15)
fatigue <- data.frame(class, time)
```

a. The model for a completely randomized design with *t* treatments and *ni* observations per treatment is
$y_{ij} = \mu_i + \varepsilon_{ij}$ with $\mu_i = \mu + \tau_i$.

*i* = 1, 2, 3, 4

*j* = 1, 2, 3, 4, 5, 6, 7

$y_{ij}$ = fatigue time for jth individual having ith weight class

$\mu_i$ = ith treatment mean

$\mu$ = overal treatment mean

$\tau_i$ = an effect due to treatment *i*, an unknown constant

$\varepsilon_{ij}$ = a random error term

The **response variable** is time to fatigue (in minutes), and the **explanatory factor** is weight class.

b.
```{r}
# 1-way ANOVA
m2 = lm(time ~ class, data = fatigue)
anova(m2)

# critical F-value
qf(.975, 3, 24)
```
The p-value corresponding to weight class is ~0, which is less than a=0.05, thus there is significant difference in the fatigue time between different weight classifications. Therefore it can be concluded that there is a statistically significant relationship between weight and exercise tolerance.

c. To design an experiment to simultaneously measure the effects of Age, Gender, Excess Weight on Fatigue time, the design will be a factorial treatment structure with three factors: A = age, B = gender, and C = excess weight. Then the design is a completely randomized a X b X c factorial experiment.

## 3. Video games revisited

1-Factor, k=4, n=40, balanced

df1 = 3, df2 = 36

a. six pairs of t-tests: (7plus-12plus, 7plus-16plus, 7plus-18plus, 12plus-16plus, 12plus-18plus, 16plus-18plus)

```{r}
## group means
aggregate(VideoGames$Desire, list(VideoGames$Label), FUN = mean)
```

a.
```{r}
library(agricolae)
m1 = lm(Desire ~ Label, data = VideoGames)
LSD <- LSD.test(m1, "Label")
LSD
```

b.
```{r}
library(agricolae)
m1aov <- aov(Desire ~ Label, data = VideoGames)
TukeyHSD(m1aov)
```
c. Differences between the models

## 5. Used cars

```{r}
usedcar <- read.table("~/Desktop/UGA/Fall 2020/STAT 6315/hw/hw 6/usedcar.txt", quote="\"", comment.char="")

# create data frame
age <- c(rep("young", 12), rep("middle", 12), rep("elder", 12))
price <- c(2300,2450,2150,2200,2100,2225,2075,2250,1975,2225,1900,2100,
           2750,2675,2700,2850,2625,2900,2725,2975,2800,2650,2600,2900,
           2350,2000,2550,2150,2175,2325,2100,2000,1950,2025,2200,2075)
car <- data.frame(age, price)

# 1-way ANOVA
m5 = lm(price ~ age)
anova(m5)
summary(m5)
```
RMSE = 150

```{r}


```






