---
title: "HW5"
author: "wes bonelli"
output: html_notebook
---

## 1. Verbal abilities

```{r}
nm = 100
nw = 100
ym = 500
yw = 550
sm = 120
sw = 110
verbal_t = qt(0.975, df = 198)
verbal_sp = sqrt(((99 * sm ^ 2) + (99 * sw ^ 2)) / 198)
verbal_lb = ym - yw - verbal_t * verbal_sp * sqrt(1/100 + 1/100)
verbal_ub = ym - yw + verbal_t * verbal_sp * sqrt(1/100 + 1/100)

verbal_lb
verbal_ub
```

The 95% confidence interval is approx. [-82.1, -17.9].

## 2. Verbal/Math Abilities

The confidence interval cannot be found. To find a paired sample confidence interval, it would be necessary to know the dependence between samples in order to estimate standard error.

## 3. NAEP

```{r}
n = 400
p1 = 0.85
q1 = 1 - p1
p2 = 0.77
q2 = 1 - p2
alpha = 0.13
z = qnorm(1 - (alpha / 2))
mu1 = n * p1
mu2 = n * p2
sigma1 = sqrt(n * p1 * q1)
sigma2 = sqrt(n * p2 * q2)
diff = mu1 - mu2
se_diff = sqrt((sigma1 ^ 2) / n + (sigma2 ^ 2) / n)
lb = diff - z * se_diff
percent_lb = lb / 400 * 100
ub = diff + z * se_diff
percent_ub = ub / 400 * 100

percent_lb
percent_ub
```

Using the normal approximation to the binomial, the 87% confidence interval for the true percentage drop is approx. [7.79%, 8.21%]. It's highly likely that there was a drop.

## 4. Dissolved Oxygen

### 4a.

Just to sanity check, find the t-statistic the hard way:

```{r}
n = 15
y1 = c(5.2, 4.8, 5.1, 5.0, 4.9, 4.8, 5.0, 4.7, 4.7, 5.0, 4.6, 5.2, 5.0, 4.9, 4.7)
y2 = c(3.2, 3.4, 3.7, 3.9, 3.6, 3.8, 3.9, 3.6, 4.1, 3.3, 4.5, 3.7, 3.9, 3.8, 3.7)
m_y1 = mean(y1)
m_y2 = mean(y2)
diff = m_y1 - m_y2
s_y1 = sd(y1)
s_y2 = sd(y2)
s_p = sqrt(((14 * s_y1 ^ 2) + (14 * s_y2 ^ 2)) / 28)
t = (diff - 0) / (s_p * sqrt(1 / n + 1 / n))
```

Perform the t-test:

```{r}
y1 = c(5.2, 4.8, 5.1, 5.0, 4.9, 4.8, 5.0, 4.7, 4.7, 5.0, 4.6, 5.2, 5.0, 4.9, 4.7)
y2 = c(3.2, 3.4, 3.7, 3.9, 3.6, 3.8, 3.9, 3.6, 4.1, 3.3, 4.5, 3.7, 3.9, 3.8, 3.7)
t.test(y1, y2, alternative = "greater", conf.level = 0.99)
```

The data do provide evidence that there was a significant reduction in mean dissolved O.

### 4b.

```{r}
hist(y1)
hist(y2)
```

Neither dataset looks particularly normal, but that does not preclude the t-test. It is unclear whether true random sampling was used, but if one assumes so (and that samples were independent) then the conditions for the t-test are satisfied.

### 4c.

The significance level (p-value) is 1.065e-11.

### 4d.

```{r}
t.test(y1, y2, alternative = "two.sided", conf.level = .99)
```

The 99% confidence interval for the true drop in mean dissolved O is approx. [0.9, 1.44].

## 5. Twins

```{r}
library("readxl")
twins = read_excel("Twins.xls")
```


### 5a.

```{r}
t.test(twins$Acad, twins$NonA, conf.level = 0.95, paired = TRUE)
```

Yes, there is likely a difference in the mean final grades (p-value = 2.918e-05).

### 5b.

The 95% confidence interval for size of the difference in mean final grades is approx. [2.23, 5.37].

### 5c.

```{r}
hist(twins$Acad)
hist(twins$NonA)
```

Both datasets look approximately normal. If the twin pairs were randomly sampled, then the conditions for the paired samples t-test seem to be satisfied.

### 5d.

???

## 6. Economists

```{r}
n1 = 92
y1 = 5.75
s1 = 1.2
n2 = 93
y2 = 5.27
s2 = 1.5
```

### 6a.

```{r}

```

